---
title: "Human height, weight and age"
author: "Pablo Espeso"
date: "6/6/2020"
output: html_document
---

## Submission

This submission contains two files:

* [html file](plotly_assignment_doc.html) (this file!), with an explanation of the assignment and the created project.
* [Rmd source file](plotly_assignment_doc.Rmd), with the code of the html file.

## Assignment requeriments

_From Coursera_

Create a web page presentation using R Markdown that features a plot created with Plotly. Host your webpage on either GitHub Pages, RPubs, or NeoCities. Your webpage must contain the date that you created the document, and it must contain a plot created with Plotly. We would love to see you show off your creativity!

Criteria:

1. Does the web page feature a date and is this date less than two months before the date that you're grading this assignment?
2. Is the web page a presentation and does it feature an interactive plot that appears to have been created with Plotly?

## Data explanation

We took a dataset with height (meters), weight (kilograms), age (years) and gender (1 == male; 0 == female) and plot all of them in a scatter3d plotly.

Data is taken from [Howell1.csv](https://github.com/rmcelreath/rethinking/blob/master/data/Howell1.csv) and can be downloaded [from here](https://github.com/rmcelreath/rethinking/raw/master/data/Howell1.csv). It contains 545 rows with all data. An aalternative dataset is [Howell2](https://github.com/rmcelreath/rethinking/blob/master/data/Howell2.csv), slightly different.

For this simple assignment, we take the data directly from the csv and show the scatter3d with these variables:

* X: height
* Y: weight
* Z: age

And color them with the gender. Since it is a binary variable (1 == male; 0 == female), the plot only shows two colours.

## Code explanation

### Load libraries

First, we load the plotly library.

```{r libraries, message = FALSE}
library(plotly)
```

### Load data

We set a url to download the data and load it into a `data` variable.


```{r load_data}
url <- "https://raw.githubusercontent.com/rmcelreath/rethinking/master/data/Howell1.csv"
data <- read.csv(url, sep = ";")
str(data)
```

### Create and show plotly

Once the data has been loaded, we create a plotly plot and show it.

```{r plot_ly, warning = FALSE}
my.plot <- plot_ly(data,
                   type = "scatter3d",
                   mode = "markers",
                   x = ~height,
                   y = ~weight,
                   z = ~age,
                   color = ~male)
my.plot <- my.plot %>% layout(scene = list(
              xaxis = list(title = "Height (meters)"),
              yaxis = list(title = "Weight (kilograms)"),
              zaxis = list(title = "Age (years)")
              )
)
my.plot

```